Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2021-09-14 RR_manuscript/Draft_RR.Rmd
include reference to recruitment form and analytical approach questionnaire, includes temp. links

---
title: "Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses"
runninghead: "Multidimensional signals and analytic flexibility"
author:
- name: Stefano Coretta*
  num: 1
- name: Joseph V. Casillas
  num: 2
- name: Timo B. Roettger
  num: 3
address:
- num: 1
  org: Institute of Phonetics and Speech Processing, Ludwig-Maximilian University Munich, Germany.
- num: 2
  org: Department of Spanish and Portuguese, Rutgers University, New Brunswick, United States
- num: 3
  org: Department of Linguistics and Scandinavian Studies, University of Oslo, Oslo, Norway
corrauth: "Stefano Coretta"
email: "s.coretta@lmu.de"
abstract: "Recent efforts to replicate published findings have uncovered surprisingly low success rates across disciplines. Moreover, several studies have highlighted the large degree of analytic flexibility in data analysis which can lead to substantially different conclusions based on the same data set. Thus, researchers have expressed their concerns that these researcher degrees of freedom might facilitate bias and can lead to claims that do not stand the test of time. Even greater flexibility is to be expected in fields in which the primary data lend themselves to a variety of possible operationalizations. The multidimensional, temporally extended nature of speech constitutes an ideal testing ground for assessing the variability in analytic approaches that derives not only from aspects of statistical modelling but also from decisions regarding the quantification of the measured behavior. In the present study, we will give the same speech production dataset to N teams of researchers and will ask them to answer the same research question. Using Bayesian meta analytical tools, we will quantify this variability and relate it to predictors related to the analysts and their analyses."
keywords: "crowdsourcing science, data analysis, scientific transparency, speech, acoustic analysis"
classoption:
  - Review
  - times
bibliography: references
bibliographystyle: sageh
output:
  bookdown::pdf_book:
    base_format: rticles::sage_article
    keep_tex: yes
    template: template.tex
    includes:
      in_header: preamble.tex
---

@@

In this article, we investigate the impact of analytic choices on research results when many analyst teams analyze the same speech production data set, a process that involves both decisions regarding the *operationalization* of a complex signal and decisions regarding *statistical analysis*.
<!-- TODOLATER: include number of teams when known -->
Specifically, we discuss the degree of variability in research results obtained by N teams who had to choose the operationalisation and statistical procedures to answer the same research question, on the basis of the same same set of raw data (here, speech recordings).
Our goals are twofold: First, our study conceptually replicates previous many-analyses projects, by probing the effects of different statistical analyses and by assessing the generalizability of published findings to other disciplines (here, the speech sciences).
Second, our study extends the scope of inquiry to include flexibility in the operationalisation of complex human behavior (here, human speech).
This is an important addition in that the increased number of "forking paths" in the "garden of analytic choices", deriving from the many decisions involved in quantification, might reveal a higher degree of variability across analysts than what previously observed, thus giving us a more realistic estimate of variability across analysts.

@@

Referring is one of the most basic and prevalent uses of language and one of the most widely researched areas in language science. 
When trying to refer to a banana, what does a speaker say and how do they say it in a given context.
The context within which an entity occurs (i.e., with other non-fruits, other fruits, or other bananas) plays a large part in determining the choice of referring expressions.
Generally, speakers aim to be as informative as possible to uniquely establish reference to the intended object, but they are also resource-efficient in that they avoid redundancy [@grice1975logic].
Thus one would expect the use of a modifier, for example, only if it is necessary for disambiguation.
For instance, one might use the adjective *yellow* to describe a banana in a situation in which there is a yellow and a less ripe green banana available, but not when there is only one banana to begin with.

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
Focused units denote the set of all (contextually relevant) alternatives [e.g. @rooth1992theory].
Concretely, a focus contrast marks one of more elements in a sentence as prominent, by different linguistic means depending on the language.
For instance, if the competitor and target objects differ but their color did not (e.g. *yellow banana* vs *yellow tomato*), the noun is in focus (the Noun Focus condition, NF).
If the objects are the same but differ in color (e.g. *yellow banana* vs *blue banana*), the color adjective is in focus (the Adjective Focus condition, AF).
If both the color and the object differ (e.g. *yellow banana* vs *blue tomato*), the the whole noun phrase is in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
Crucially, the color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g. *orange mandarin*), medium typical (e.g. *green tomato*), or atypical (e.g. *yellow cherry*), as established by a norming study that was conducted prior to the production experiment just described.^[A detailed description of the norming and production studies from the *Prosodic encoding of redundant referring expressions* project, which was given to the analysts with the data set, can be found in `methods_norm_prod.pdf` at (temporary link) <https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f>]
Each subject produced 15 critical trials (NF condition).
Each trial was repeated twice, yielding a total of 30 trials per participant and a grand-total of 900 ($15 \times 2 \times 30$ participants) spoken utterances.

@@

We estimate that this process, from the time of an in-principle acceptance of this Stage 1 Registered Report to the end of Phase 5, will take nine months.
The factor most likely to delay our time-line is the rate of completion of the original set of analyses by independent analysis teams.
The following sections report the criteria for sample size, data exclusions, data manipulations, and all the measures in the study.

@@

An online landing page will provide a general description of the project, including a short pre-recorded slide-show that summarizes the data set and research question (<https://many-speech-analyses.github.io>).^["Questions" and "Join Project" buttons will be functioning when recruitment starts. "Join Project" will link to the intake form.]
The project will be advertised via social media, using mailing lists for linguistic and psychological societies, and via word of mouth.
The target population comprises active speech science researchers with a graduate/doctoral degree (or currently studying for a graduate/doctoral degree) in relevant disciplines.
Researchers can choose to work independently or in a small team.
For the sake of simplicity, we will refer both to a single researcher and teams as \textsc{analysis teams}.^[Terms in small caps in this and later sections are included with their definition in the glossary at the end of the paper for the reader's convenience.]
Recruitment for this project will commence upon receiving in-principle acceptance.
As outlined above, our primary aim is to assess the variability of the reported effects, rather than the meta-analytic estimate of the investigated effect *per se*.
To estimate the degree of uncertainty around effect variability as driven by number of teams, we ran a series of sample size simulations with values of variability extracted from @silberzahn2018many.
Variability among teams was operationalized as the standard deviation of the teams' reported effects from @silberzahn2018many (which we *z*-scored prior to simulations to make it comparable to our study).
For the mean of the teams' true standard deviation (0.68 *z*-score), the simulation indicates that the degree of uncertainty around the estimated teams' standard deviation will be below 1 SD at any sample size greater than 10 teams. 
Thus in order to achieve our main goal, i.e. estimating variability among teams, we consider a minimum sample size of 10 teams as sufficient. Given the exploratory nature of our study, however, we will sample as many analysts as possible. We have received initial expression of interest to participate from more than 200 analysts.

After submitting their analyses, we will ask analysts to also function as peer-reviewers. Each team must review four other analyses.

All analysts will share co-authorship on this manuscript and will participate in the collaborative process of producing the final manuscript.
Informed consent will be obtained as part of the intake form.

@@

The analysis teams will register for participation and each of the analysts individually will answer a demographic and expertise questionnaire (`intake_form.pdf`).
PDF versions of all the questionnaires used in this study are available at LINK^[Temporary link: <https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f>, Questionnaires component.].
The questionnaire collects information on the analysts current position and self-estimated breadth and level of statistical expertise and acoustic analysis skills.
We will then request that they answer the question: *Do speakers acoustically modify utterances to signal atypical word combinations?*
To do so, they will be given the data generated by the experiment described in Section [The data set](#s:dataset).
The data will include the audio recordings with corresponding time-aligned transcriptions in the form of Praat TextGrid files (LINK^[Temporary link: <https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f>, Data component.]).

@@

```{r sdi, eval=FALSE, include=FALSE}
x <- c("a", "b", "c", "d")
y <- c("b", "d", "e", "f")
(2 * length(intersect(x,y))) / (length(x) + length(y))
```

@@

All relevant data, code, and materials will be publicly archived on the Open Science Framework (<https://osf.io/3bmcp/>).^[Link for peer-review: <https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f>.]
Archived data will include the original data set distributed to all analysts, any edited versions of the data analyzed by individual teams, and the data we analyze with our meta-analyses, which include the standardized effect sizes, the statistics describing variation in model structure among analysis teams, and the anonymized answers to our questionnaires of analysts.
Similarly, we will archive both the analysis code used for each individual analysis and the code from our meta-analyses.
We will also archive copies of our survey instruments from analysts and peer-reviewers.

>>>>

---
title: "Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses"
runninghead: "Multidimensional signals and analytic flexibility"
author:
- name: Stefano Coretta*
  num: 1
- name: Joseph V. Casillas
  num: 2
- name: Timo B. Roettger
  num: 3
address:
- num: 1
  org: Institute of Phonetics and Speech Processing, Ludwig-Maximilian University Munich, Germany.
- num: 2
  org: Department of Spanish and Portuguese, Rutgers University, New Brunswick, United States
- num: 3
  org: Department of Linguistics and Scandinavian Studies, University of Oslo, Oslo, Norway
corrauth: "Stefano Coretta"
email: "s.coretta@lmu.de"
abstract: "Recent efforts to replicate published findings have uncovered surprisingly low success rates across disciplines. Moreover, several studies have highlighted the large degree of analytic flexibility in data analysis which can lead to substantially different conclusions based on the same data set. Thus, researchers have expressed their concerns that these researcher degrees of freedom might facilitate bias and can lead to claims that do not stand the test of time. Even greater flexibility is to be expected in fields in which the primary data lend themselves to a variety of possible operationalizations. The multidimensional, temporally extended nature of speech constitutes an ideal testing ground for assessing the variability in analytic approaches that derives not only from aspects of statistical modelling but also from decisions regarding the quantification of the measured behavior. In the present study, we will give the same speech production dataset to N teams of researchers and will ask them to answer the same research question. Using Bayesian meta-analytic tools, we will quantify this variability and relate it to predictors related to the analysts and their analyses."
keywords: "crowdsourcing science, data analysis, scientific transparency, speech, acoustic analysis"
classoption:
  - Review
  - times
bibliography: references
bibliographystyle: sageh
output:
  bookdown::pdf_book:
    base_format: rticles::sage_article
    keep_tex: yes
    template: template.tex
    includes:
      in_header: preamble.tex
---

@@

In this article, we investigate the impact of analytic choices on research results when many analyst teams analyze the same speech production data set, a process that involves both decisions regarding the *operationalization* of a complex signal and decisions regarding *statistical analysis*.
<!-- TODOLATER: include number of teams when known -->
Specifically, we discuss the degree of variability in research results obtained by N teams who had to choose the operationalisation and statistical procedures to answer the same research question, on the basis of the same set of raw data (here, speech recordings).
Our goals are twofold: (i) our study conceptually replicates previous many-analyses projects, by probing the effects of different statistical analyses and by assessing the generalizability of published findings to other disciplines (here, the speech sciences);
(ii) our study extends the scope of inquiry to include flexibility in the operationalisation of complex human behavior (here, human speech).
This is an important addition in that the increased number of "forking paths" in the "garden of analytic choices", deriving from the many decisions involved in quantification, might reveal a higher degree of variability across analysts than what previously observed, thus giving us a more realistic estimate of variability across analysts.

@@

Referring is one of the most basic and prevalent uses of language and one of the most widely researched areas in language science. 
When trying to refer to a banana, what does a speaker say and how do they say it in a given context.
The context within which an entity occurs (i.e., with other non-fruits, other fruits, or other bananas) plays a large part in determining the choice of referring expressions.
Generally, speakers aim to be as informative as possible to uniquely establish reference to the intended object, but they are also resource-efficient in that they avoid redundancy [@grice1975logic].
Thus one would expect the use of a modifier, for example, only if it is necessary for disambiguation.
For instance, one might use the adjective *yellow* to describe a banana in a situation in which there are a yellow and a less ripe green banana available, but not when there is only one banana to begin with.

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
Focused units denote the set of all (contextually relevant) alternatives [e.g. @rooth1992theory].
Concretely, a focus contrast marks one of more elements in a sentence as prominent, by different linguistic means depending on the language.
For instance, if the competitor and target objects differ but their color does not (e.g. *yellow banana* vs *yellow tomato*), the noun is said to be in focus.
We call this the Noun Focus condition, or NF for short.
If the objects are the same but differ in color (e.g. *yellow banana* vs *blue banana*), the color adjective is in focus (the Adjective Focus condition, AF).
If both the color and the object differ (e.g. *yellow banana* vs *blue tomato*), then the whole noun phrase is in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
Crucially, the color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g. *orange mandarin*), medium typical (e.g. *green tomato*), or atypical (e.g. *yellow cherry*), as established by a norming study that was conducted prior to the production experiment just described.^[A detailed description of the norming and production studies from the *Prosodic encoding of redundant referring expressions* project, which was given to the analysts with the data set, can be found in `methods_norm_prod.pdf` at (temporary link) <https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f>]
Each subject produced 15 critical trials (NF condition).
Each trial was repeated twice, yielding a total of 30 trials per participant and a grand-total of 900 ($15 \times 2 \times 30$ participants) spoken utterances.

@@

We estimate that this process, from the time of an in-principle acceptance of this Stage 1 Registered Report to the end of Phase 5, will take nine months.
The factor most likely to delay our time-line is the rate of completion of the original set of analyses by independent analysis teams.

The project OSF repository contains all the materials mentioned in this paper and can be accessed at <https://bit.ly/3AqU1ul>.
The main repository holds three OSF components, `Data`, `Questionnaires`, and `Cache`, and it is linked to the GitHub repository.
The following sections report the criteria for sample size, data exclusions, data manipulations, and all the measures in the study.

@@

An online landing page will provide a general description of the project, including a short pre-recorded slide-show that summarizes the data set and research question (<https://many-speech-analyses.github.io>).^["Questions" and "Join Project" buttons will be functioning when recruitment starts. "Join Project" will link to the intake form.]
The project will be advertised via social media, using mailing lists for linguistic and psychological societies, and via word of mouth. 
Social media advertising will be accompanied by a short recruitment form available at LINK.^[Temporary link: <https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f>, Recruitment form.]
The target population comprises active speech science researchers with a graduate/doctoral degree (or currently studying for a graduate/doctoral degree) in relevant disciplines. 
All individuals interested in participating will be asked to complete a questionnaire detailing their familiarity with numerous analytical approaches common in the speech sciences (LINK^[Temporary link: <https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f>, Analytical approach form.]).
Researchers can choose to work independently or in a small team.
For the sake of simplicity, we will refer both to a single researcher and teams as \textsc{analysis teams}.^[Terms in small caps in this and later sections are included with their definition in the glossary at the end of the paper for the reader's convenience.]
Recruitment for this project will commence upon receiving in-principle acceptance.
As outlined above, our primary aim is to assess the variability of the reported effects, rather than the meta-analytic estimate of the investigated effect *per se*.
To estimate the degree of uncertainty around effect variability as driven by number of teams, we ran a series of sample size simulations with values of variability extracted from @silberzahn2018many.
Variability among teams was operationalized as the standard deviation of the teams' reported effects from @silberzahn2018many (which we *z*-scored prior to simulations to make it comparable to our study).
For the mean of the teams' true standard deviation (0.68 *z*-score), the simulation indicates that the degree of uncertainty around the estimated teams' standard deviation will be below 1 SD at any sample size greater than 10 teams. 
Thus in order to achieve our main goal, i.e. estimating variability among teams, we consider a minimum sample size of 10 teams as sufficient. Given the exploratory nature of our study, however, we will sample as many analysts as possible. We have received initial expression of interest to participate from more than 200 analysts.

After submitting their analyses, we will ask analysts to also function as peer-reviewers. Each team must review four other analyses.
All analysts will share co-authorship on this manuscript and will participate in the collaborative process of producing the final manuscript.
Informed consent will be obtained as part of the intake form.

@@

The analysis teams will register for participation and each of the analysts individually will answer a demographic and expertise questionnaire (`intake_form.pdf`).
A PDF version of this and all other questionnaires are available in the repository's `Questionnaires` component, at <https://bit.ly/3Cf3fdm>.
The questionnaire collects information on the analysts current position and self-estimated breadth and level of statistical expertise and acoustic analysis skills.
We will then request that they answer the question: *Do speakers acoustically modify utterances to signal atypical word combinations?*
To do so, they will be given the data generated by the experiment described in Section [The data set](#s:dataset).
The data will include the audio recordings with corresponding time-aligned transcriptions in the form of Praat TextGrid files.
These can be found in the `Data` component at <https://bit.ly/3Ahawc7>.

@@

```{r sdi, eval=FALSE, include=FALSE}
x <- c("a", "b", "c", "d")
y <- c("b", "d", "e", "f")
(2 * length(intersect(x, y))) / (length(x) + length(y))
```

@@

<!-- Had to remove footnote because there were too many footnotes and not enough symbols -->

All relevant data, code, and materials will be publicly archived on the Open Science Framework (<https://osf.io/3bmcp/>, link for peer-review: <https://bit.ly/3AqU1ul>).
Archived data will include the original data set distributed to all analysts, any edited versions of the data analyzed by individual teams, and the data we analyze with our meta-analyses, which include the standardized effect sizes, the statistics describing variation in model structure among analysis teams, and the anonymized answers to our questionnaires of analysts.
Similarly, we will archive both the analysis code used for each individual analysis and the code from our meta-analyses.
We will also archive copies of our survey instruments from analysts and peer-reviewers.
